`respR` - An R package for the analysis of respirometry data

### 2017-10-18 manuscript/manuscript-HTML.Rmd
NC - a few additional edits

Recently, a number of open-source R software packages have become available (e.g. `respirometry`: https://CRAN.R-project.org/package=respirometry, `rMR`: https://CRAN.R-project.org/package=rMR, `LoLinR`: Olito et al., 2017) which are suitable for or designed to analyse respirometry data. ~~ Both `respirometry` and `rMR` are useful for aquatic respirometry, but are not focused on the analysis of data, although both packages have functions for the analysis of metabolic rate over non-overlapping intervals, and `rMR` has a specialised function to calculate $P_{c}$. ~~ [nc - need a better way of basically saying, they're ok, but very limited, and not user-friendly, or adaptable, or]. `LoLinR` provides a very statistically robust method of detecting a "best-fit" regression which is applicable for closed and flow-through respirometry. [nc - try to add something else positive about it! We need to be nice to them]. However, `LoLinR`'s processing time increases exponentially with large datasets, with data containing more than 500 observations taking several minutes to run (Olito et al., 2017), which makes `LoLinR` somewhat intractable with large amounts of data. In addition, its focus is on data which is evenly spaced with regard to time, and where the response variable changes in a monotonic fashion, which is far from typical of respirometry data.

Here we describe the R package `respR`, a set of functions desgined to provide a structured, reproduceable workflow for the analysis of respirometry-related data (Fig. 1). The package contains utilities to: (1) analyse closed, intermittent, flow-through and open respirometry data, (2) scale $\Delta O_2$ to volume and/or mass, and surface area [nc - soon yes?] (3) automatically detect maximal and minimal rates, $P_{c}$, or the most linear results in large datasets with the help of traditional rolling regression and kernel density estimation techniques. Other functions useful in processing respirometry data are avaliable, and are described in more detail in our online html vignette (https://januarharianto.github.io/respR/). While the package has a strong focus on aquatic respirometry, similar principals apply to aerial respirometry. ~~Some oxygen concentration conversions require a salinity value, but in freshwater respirometry this should be set to zero.~~ [nc - don't need this, can go in the help docs]. We demonstrate the utility of the package by analysing three datasets as case studies and proof of concept. These data were collected using closed, intermittent and flow-through respirometry.

[nc - i would drop closed, focus on intermittent (which is same as closed essentially) and flowthrough, and include a PCrit case. It's a major selling point! Several people i have told about it have asked - does it do PCrit?]

>>>>

Recently, a number of open-source R software packages have become available (e.g. `respirometry`: https://CRAN.R-project.org/package=respirometry, `rMR`: https://CRAN.R-project.org/package=rMR, `LoLinR`: Olito et al., 2017) which are suitable for or designed to analyse respirometry data. ~~Both `respirometry` and `rMR` are useful for aquatic respirometry, but are not focused on the analysis of data, although both packages have functions for the analysis of metabolic rate over non-overlapping intervals, and `rMR` has a specialised function to calculate $P_{c}$.~~ [nc - need a better way of basically saying, they're ok, but very limited, and not user-friendly, or adaptable, or]. `LoLinR` provides a very statistically robust method of detecting a "best-fit" regression which is applicable for closed and flow-through respirometry. [nc - try to add something else positive about it! We need to be nice to them]. However, `LoLinR`'s processing time increases exponentially with large datasets, with data containing more than 500 observations taking several minutes to run (Olito et al., 2017), which makes `LoLinR` somewhat intractable with large amounts of data. In addition, its focus is on data which is evenly spaced with regard to time, and where the response variable changes in a monotonic fashion, which is far from typical of respirometry data.

Here we describe the R package `respR`, a set of functions desgined to provide a fast [or better word], structured, reproduceable workflow for the analysis of respirometry-related data (Fig. 1). The package contains utilities to: (1) analyse closed, intermittent, flow-through and open respirometry data, (2) determine volume-, mass-, and surface area- [nc - soon yes?] specific oxygen uptake rates, (3) automatically detect maximal and minimal rates, $P_{c}$, or the most linear results in large datasets with the help of traditional rolling regression and kernel density estimation techniques. Other functions useful in processing respirometry data are avaliable, and are described in more detail in our online html vignette (https://januarharianto.github.io/respR/). The package has a strong focus on aquatic respirometry, however similar principals apply to aerial respirometry, so the analysis workflow should also be suitable for these data. ~~Some oxygen concentration conversions require a salinity value, but in freshwater respirometry this should be set to zero.~~ [nc - don't need this, can go in the help docs]. We demonstrate the utility of the package by analysing three datasets as case studies and proof of concept. These data were collected using closed, intermittent and flow-through respirometry.

[nc - i would drop closed, focus on intermittent (which is same as closed essentially) and flowthrough, and include a PCrit case. It's a major selling point! Several people i have told about it have asked - does it do PCrit? Or i guess, make the closed e.g. the PCrit one... ]
